Learning action models with minimal observability

### 2019-05-01 doc/aij18/minor-revision/background.tex
minor modifications

Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a comparison between the learned model and a {\em Ground-Truth Model}), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).

>>>>

Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2}), the metrics used in the evaluation (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}). \textcolor[rgb]{1.00,0.00,0.00}{Regarding the evaluation method, almost all approaches rely on a comparison between the learned model and a {\em Ground-Truth Model} (GTM). By GTM we refer to an engineered model that knowledge modelers have adopted as being correct with respect to a domain, and that the planning community accepts as such.}

### 2019-05-01 doc/aij18/minor-revision/evaluation.tex
minor modifications

We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model. This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM. The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models. To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}. These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:

>>>>

\textcolor[rgb]{1.00,0.00,0.00}{As commented before, the large majority of approaches shown in Table \ref{table:models_comparison2} evaluate the learned action models wrt GTM, where the GTM is used to generate the test set of plan traces that the algorithms use to learn from. That is, learning approaches are evaluated in domains taken from the IPCs.} We can also observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model. This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM. The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models. To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}. These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:

### 2019-05-01 doc/aij18/minor-revision/learning_action_models.tex
minor modifications

When the input plan trace contains observed actions, the extra conditional effects
$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$. \textcolor{red}{As a side note, if we are sure that the input trace contains all actions (\FO sequence of actions) this conditional effect should be redefined as regular preconditions and effects. Doing so will make it impossible for the planner to execute apply actions outside of the observed ones. Additionally, for traces with \POstar state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\

>>>>

When the input plan trace contains observed actions, the extra conditional effects
$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$. \textcolor{red}{As a side note, when the input trace contains all actions (\FO sequence of actions), this conditional effect should be redefined as regular preconditions and effects. Doing so will make it impossible for the planner to execute apply actions outside of the observed ones. Additionally, for traces with \POstar state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\

### 2019-05-01 doc/aij18/minor-revision/learning_task_def.tex
minor modifications

\item \textbf{$\tau$ does not identify the length of $\pi$}. This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories. In this case, we are unaware of the number of actions of $\pi$ and the number of states induced by $\pi$. This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so \textbf{the horizon of $\pi$ is no longer known}. This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$. \textcolor{red}{SAT approaches, on the other hand, are no longer obviously applicable given the lack of a length bound}. In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown). Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.

>>>>

\item \textbf{$\tau$ does not identify the length of $\pi$}. This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories. In this case, we are unaware of the number of actions of $\pi$ and the number of states induced by $\pi$. This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so \textbf{the horizon of $\pi$ is no longer known}. This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$. \textcolor{red}{SAT approaches, on the other hand, are no longer obviously applicable given the lack of a length bound}. \textcolor[rgb]{0.00,0.00,1.00}{In this case, the learning task is not SAT compilable and thereby we can no longer leverage the potential of SAT solvers given the lack of a length bound.} In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown). Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.
